---
title: 'Coursera''s Statistical Inference Final Project - Part 2: Inferential Analysis'
author: "Daniel Szulc"
date: "Friday, April 24, 2015"
output:
  pdf_document: default  
---
```{r echo=FALSE}
library(knitr)
opts_chunk$set(fig.width=4, fig.height=3, dev = 'pdf')
```

## Overview

## Exploratory data analysis (basic summary of the data)
```{r exploratory}
library(datasets)
library(xtable)
        library(dplyr,warn.conflicts = FALSE)
data(ToothGrowth)
figure <- group_by(ToothGrowth,supp,dose) %>% 
        summarise(count = n(), mean.len=mean(len), 
                  sd.len=signif(sd(len), digits = 3))
figure <- xtable(figure, caption="Basic summary statistics by supp and dose")
```

```{r exploratory_table, results='asis'}
print(figure, caption.placement="top", include.rownames=FALSE, comment=FALSE)
```
```{r exploratory_plots, fig.width=8}
par(mfrow=c(1,2), mar=c(4,3,2,1))
boxplot(len~supp, data=ToothGrowth, col="yellow", main="Tooth Growth by Delivery Method")
boxplot(len~dose, data= ToothGrowth, col="orange", main="Tooth Growth by Dose Level")

```
```{r exploratory_plot2, fig.width=8, fig.cap="Fig"}
library(ggplot2)
ggplot(ToothGrowth,aes(len))+geom_density(size=1)+theme_bw()+facet_grid(supp~dose)+geom_vline(xintercept=mean(ToothGrowth$len), col="red",linetype=2, size=0.8)+labs(title="Tooth Growth by Delivery Method and Dose")

```


## Hypothesis test 

### 1. Is there a difference between the tooth length by supp(OJ,VC)?
$H_0$: There is no difference between supp OJ and VC in terms of tooth length.  
$H_A$: There is a difference between supp OJ and VC in terms of tooth length.  

```{r ht_1}

t.test(len ~ supp, data=ToothGrowth, paired = FALSE)
```

**As the p-value is >0.05 we can conclude: no convincing evidence supporting hypothesis that there is a difference between OJ and VC**

### 2. Is there a difference between the tooth length by dose(0.5,1,2)?

$H_0$: There is no difference between dose 0.5 and 1 in terms of tooth length.  
$H_A$: There is a difference between supp 0.5 and 1 in terms of tooth length.  

```{r ht_2}
library(dplyr)
ToothGrowth.filtered<-filter(ToothGrowth, dose %in% c(0.5,1))

t.test(len ~ dose, data=ToothGrowth.filtered, paired = FALSE)
```

**As the p-value is <0.05 we can conclude: there is convincing evidence supporting hypothesis that there is a difference between 0.5 and 1 dose**

$H_0$: There is no difference between dose 0.5 and 1 in terms of tooth length.  
$H_A$: There is a difference between supp 0.5 and 1 in terms of tooth length.  

```{r ht_3}
library(dplyr)
ToothGrowth.filtered<-filter(ToothGrowth, dose %in% c(1,2))

test.3 <- t.test(len ~ dose, data=ToothGrowth.filtered, paired = FALSE)
print(test.3)

```

**As the p-value is <0.05 we can conclude: there is convincing evidence supporting hypothesis that there is a difference between 1 and 2 dose**

Given the results of the above test and taking into account basic summary of data (particularly the means in groups), we can be pretty sure that there is also a significant difference between dose 0.5 and 2.

## Conclusions

