---
title: 'Coursera''s Statistical Inference Final Project - Part 1: Simulation'
author: "Daniel Szulc"
date: "Friday, April 24, 2015"
output:
  pdf_document: default  
---
```{r echo=FALSE}
library(knitr)
opts_chunk$set(fig.width=8, fig.height=3, dev = 'pdf')
```
# Report
## Overview

The aim of this report is to present a comparison between the simulated exponential distribution (using R random number generator) and theoretical distribution of sample means from exponential distribution (calculated based on Central Limit Theorem).
We are going to see if the simulated sampling distribution is in line (in terms of centre, spread and shape) with what we would expected based on Central Limit Theorem.

## Simulation

Let's set a random number generator seed (to make sure that this simulation is fully reproducible) and simulate the exponential distribution.  
We are conducting two simulations.

```{r}
set.seed(seed = 10)
n <- 1000
lambda <- 0.2
sample.size <- 40

population <- rexp(n,lambda)
means <- sapply(seq_len(n), function(x) mean(rexp(sample.size,lambda)))
```

We are conducting two simulations. In the first we are drawing 1000 observations from the exponentially distributed population (with $\lambda=0.2$). The second simulation drawing 1000 samples from the same population. The size of each sample is 40. The plot below show the effects of two simulations. The observations from the population (the first simulation) is plotted with light green colour. The distribution of sample means (the second simulation) is plotted with "salmon" colour.

```{r}
library(ggplot2)
ggplot(data.frame(x=c(population,means),
                  label=factor(rep(c("population","means"),c(n,n)))),
       aes(x, fill=label))+
geom_density(alpha=0.2)
```

## Sample Mean versus Theoretical Mean
According to the Central Limit Theorem, we expect that distribution of sample means for large enough sample size is centred at true population mean. We know that the mean of exponential distribution is equal $\frac{1}{\lambda}$, therefore in this case $\frac{1}{0.2}=5$ (red vertical line in the plot below). Simulated sample mean in our case is equal to `r mean(means)` (blue line in the plot below). As we can see, it is very close to the CLT theoretical mean. 

```{r}
library(ggplot2)
ggplot(data.frame(x=means), aes(x))+geom_density(alpha=0.2, fill="salmon")+
        geom_vline(xintercept=mean(means),color="blue")+
        geom_vline(xintercept=1/lambda, color="red")+
        labs(title="Sample mean (sample size=40) versus Theoretical Mean", 
             size=3)+
        annotate("text", label = "Theoretical mean= 1/lambda", x = 3.1, y = 0.4,
                 size = 4, colour = "red", hjust=0)+
        annotate("text", label = "Sample Mean", x = 5.2, y = 0.3, size = 4, 
                 colour = "blue", hjust=0)
```
 
## Sample Variance versus Theoretical Variance

According to the Central Limit Theorem, we expect that the standard deviation of the distribution of sample means for large enough sample size is equal to population standard deviation divided by square root of sample size. We know that the standard deviation of exponential distribution is equal to $\frac{1}{\lambda}$, therefore in this case $\frac{1}{0.2}=5$. The theoretical sample standard deviation should be therefore equal to $\frac{5}{\sqrt{40}}$. As the standard deviation is the square root of the variance we know that theoretical sample variance is equal to $(\frac{5}{\sqrt{40}})^2$ so in this case `r (5/sqrt(40))^2`. The simulated sample variance is equal to `r var(means)` so it's very close.

```{r, fig.cap="Fig 2"}
library(xtable)
figure <- xtable(
        data.frame(Description=c("Theoretical Variance","Sample Variance"), 
                   Value=c((1/lambda/sqrt(sample.size))^2,var(means))),
        caption="Comparision Theoretical Variance versus Sample Variance")
```
```{r results='asis'}
print(figure, comment=FALSE, caption.placement="top", include.rownames=FALSE)
```

## Approximately normal sampling distribution
According to the Central Limit Theorem, we expect that the distribution of sample means is normal for large enough sample sizes. The plot below presents the distribution of observations drawn from the normal distribution with the mean 5 and standard deviation 5. As we can see, our distribution of sample means follow the normal distribution pattern pretty closely. 
As the additional prove please see the normal probability plot in the **Appendix**. In case of normally distributed data, all data points should lay close to the diagonal line. We can see that this condition is fairly met. Minor deviations in the both end of distribution may be due to not large enough sample size (remember, it is only 40 observations).

```{r fig.width=8}
ggplot(
        data.frame(x=c(rnorm(n,1/lambda,1/lambda/sqrt(sample.size)),means), 
                   label=factor(rep(c("normal distribution","distribution of means"),c(n,n)))), 
       aes(x, fill=label))+geom_density(alpha=0.2)

```

# Appendix
```{r}
qqnorm(means)
qqline(means)

```